---
title: "Growth Curves"
author: "Jill Hagey and Naomi Almanzor"
output: 
  html_document:
    theme: spacelab 
    toc: true
    toc_depth: 2
    toc_float: true
    df_print: paged
    highlight: espresso
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_knit$set(root.dir = "C:/Users/jvhagey/Desktop/") #'relative_path_to_root_from_Rmd
knitr::opts_knit$set(root.dir = "C:/Users/Jill/OneDrive - UC Davis/Documents/Disseration Lab work/Nitrogen_fixing/")
getwd()
```

```{r warning=FALSE, include=FALSE}
#loading the libraries
library(growthcurver);packageVersion("growthcurver")
library(chron); packageVersion("chron") #to fix time
library(ggplot2); packageVersion("ggplot2")
library(reshape2); packageVersion("reshape2")
library(tidyr); packageVersion("tidyr")
library(dplyr); packageVersion("dplyr")
library(tidyverse); packageVersion("tidyverse")
library(knitr); packageVersion("knitr")
library(kableExtra); packageVersion("kableExtra")
library(cowplot); packageVersion("cowplot")
```

Calling in the data.

```{r}
#calling in the data
d <- read.table("Growth Curves.txt", header = TRUE, sep = "\t", stringsAsFactors = FALSE)
#Fixing our times!
d$Mins <- round((d$Mins/60), 1)
d <- d[c(1:12),c(1:7)] #dropping NAs
d
```

```{r eval=TRUE, include=FALSE}
#generate names for list of column names
txs <- c("Time","7mM","3.5mM","1.75mM","0.875mM", "0.4375mM","0mM")
#adding in column names and adding back in Kinetic read times 
colnames(d) <- txs
```

First, we run the growth curve model on all the samples.

```{r eval=TRUE, include=FALSE}
#Making models for all
models.all <- lapply(d[2:ncol(d)], function(x) SummarizeGrowth(d$Time, x))
```

```{r eval=TRUE, include=FALSE}
df.predicted.plate <- data.frame(Time = d$Time)
for (i in names(d[2:ncol(d)])){
df.predicted.plate[[i]] <- predict(models.all[[i]]$model)}
#changing column name from "time" to "kinetic.read" to merge dataframe correctly
colnames(df.predicted.plate)[1] <- "Time"
#Melting dataframes to combine
melt1 <- melt(d, id.vars = "Time", variable.name = "sample", value.name = "od")
melt2 <- melt(df.predicted.plate, id.vars = "Time", variable.name = "sample", value.name = "pred.od")
df.final <- cbind(melt1, pred.od=melt2[,3])
```

##Growth Curves
Next we will graph out growth curves different concentrations with NH3.

```{r fig.height = 4, fig.width = 14, eval=TRUE, include=FALSE}
##the predict line is off since there is no blank sample. 
ggplot(df.final, aes(x=Time, y=od)) + geom_point(aes(), alpha=0.5) + geom_line(aes(y=pred.od), color="red") + facet_wrap(~sample, ncol = 12) +   theme(strip.text.y=element_text(angle=0,face = "bold"),strip.text.x=element_text(angle=0,face = "bold"), axis.text= element_text(face = "bold"), axis.title=element_text(face = "bold"), panel.grid.major.x = element_blank(),panel.grid.minor.y=element_blank())+
  labs(x="Time (hrs)", y="Optical Density")
```

Dots are the data and the red line is predicted values based on the model.

```{r fig.height = 5, fig.width = 16}
ggplot(df.final, aes(x=Time, y=od)) + geom_point(colour = "blue",aes(), alpha=0.5) + facet_wrap(~sample, ncol = 12) + geom_line(color="blue", size=1) +
  theme(strip.text.y=element_text(angle=0,face = "bold",size=11),strip.text.x=element_text(angle=0,face = "bold",size=11), 
        axis.text= element_text(face = "bold"), axis.title=element_text(face = "bold"), panel.grid.major.x = element_blank(),panel.grid.minor.y=element_blank())+
  labs(x="Time (hrs)", y="Optical Density")
```

Next we will combine the graphs of the growth curves.

```{r fig.height = 5, fig.width = 7}
ggplot(df.final, aes(x=Time, y=od, color=sample)) + geom_point(aes(), alpha=0.5) +
  geom_line(aes(x = Time, y = od, col=sample)) +
  theme_cowplot(12) +
  #theme_bw()+
  theme(strip.text.y=element_text(angle=0,face = "bold",size=11),strip.text.x=element_text(angle=0,face = "bold",size=11), 
        axis.text= element_text(face = "bold"), axis.title=element_text(face = "bold"),legend.text=element_text(face = "bold"), legend.title=element_text(face = "bold"))+ 
  labs(color= "NH3 Concentration",x="Time (hrs)", y="Optical Density")
ggsave("Nitrogen_Growth.png",device = "png", dpi=320, width = 7)
```

##Growth Stats
Calculating doubling times for each strain in different media.

```{r}
make_df <- function(data, val){
val <- val
df <- data %>% unlist(recursive = FALSE) %>% enframe() %>% unnest() %>% rename(name='NH3 Concentration', value=val)
  #print(head(df))
  #df$Media_Type <- ifelse(grepl("Hemin", df$Sample_Type), "Hemin",
  #       ifelse(grepl("Glucose", df$Sample_Type), "Glucose", "No Media Type"))
  #df$Strain <- sub("\\_.*", "", df$Sample_Type)
  #df$Nisin_Concentration <- sub(".*\\_", "", df$Sample_Type)
  return(df)
}
```


```{r}
doubling_time_all <- lapply(models.all, function(x) x$vals$t_gen)
#making kable table
doubling_time_all %>% 
  unlist(recursive = FALSE) %>%
  enframe() %>%
  #filter(str_detect(name, "ERM") | str_detect(name, "Wt")) %>%
  unnest() %>% rename(name='NH3 Concentration', value='Time (hrs)') %>%
  kable(caption="Doubling Time (hrs)") %>% 
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = F, font_size = 11) %>%
  scroll_box(width = "100%", height = "500px")
```

Calculating time of midpoint in exponential phase for each strain in different media.
```{r}
## t_mid = halfway point of exponential phase?
mid_time_all <- lapply(models.all, function(x) x$vals$t_mid)
#making kable table
mid_time_all %>% 
  unlist(recursive = FALSE) %>% 
  enframe() %>%
  #filter(str_detect(name, "ERM") | str_detect(name, "Wt")) %>%
  unnest() %>% rename(name='Media Type', value='Time (hrs)') %>%
  kable(caption="Time (hrs) at Midpoint") %>% 
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = F,font_size = 11) %>%
  scroll_box(width = "100%", height = "500px")
```

Calculating growth rate for each strain in different media.
```{r}
## t_mid = halfway point of exponential phase?
rate_all <- lapply(models.all, function(x) x$vals$r)
#making kable table
rate_all %>% 
  unlist(recursive = FALSE) %>% 
  enframe() %>%
  #filter(str_detect(name, "ERM") | str_detect(name, "Wt")) %>%
  unnest() %>% rename(name='Media Type', value='Time (hrs)') %>%
  kable(caption="Time (hrs) at Midpoint") %>% 
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = F,font_size = 11) %>%
  scroll_box(width = "100%", height = "500px")
```

```{r}
df1 <- make_df(doubling_time_all, "Doubling_Time")
df2 <- make_df(mid_time_all, "Mid_Point")
df3 <- make_df(rate_all, "Rate")
df4 <- cbind(df1, df2$Mid_Point, df3$Rate)
colnames(df4) <- c("NH3 Concentration", "Doubling Time", "Mid Point", "Rate")
df4
```
